---
title: "Stroke Project"
output:
  pdf_document: default
  html_document: default
date: "2023-12-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## LOADING DATASET 

```{r}
stroke.data <- read.csv("C:/Users/valiv/OneDrive/Desktop/Stats Proj/stroke-data.csv", na.strings="Unknown", stringsAsFactors=TRUE)
```

## VIEWING THE DATASET

```{r}
print(stroke.data)
```
## TO KNOW THE NUMBER OF ROWS AND COLUMNS

```{r}
dim(stroke.data)
```
## COLUMN NAMES

```{r}
head(stroke.data)
```
## DATA TYPE OF COLUMNS

```{r}
str(stroke.data)
```
## SUMMARY STATISTICS

```{r}
summary(stroke.data)
```
## EXPLORATORY DATA ANALYSIS(EDA)

```{r}
# column is named "stroke"
stroke_positive_count <- sum(stroke.data$stroke == 1)
stroke_negative_count <- sum(stroke.data$stroke == 0)
total_cases <- nrow(stroke.data)

stroke_positive_percentage <- (stroke_positive_count / total_cases) * 100

cat("Total Cases:", total_cases, "\n")
cat("Stroke Positive Cases:", stroke_positive_count, "\n")
cat("Stroke Negative Cases:", stroke_negative_count, "\n")
cat("Percentage of Stroke Positive Cases:", stroke_positive_percentage, "%\n")

```
## HISTOGRAM PLOT TO VIEW THE DISTRIBUTION OF BMI

```{r}
hist(stroke.data$bmi)
```
## HISTOGRAM TO VIEW THE BMI DISTRIBUTION IN STROKE POSITIVE PATIENTS
```{r}
# columns are named "bmi" and "stroke"
library(ggplot2)
ggplot(subset(stroke.data, stroke == 1), aes(x = bmi)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "BMI Distribution of Stroke-Positive Patients",
       x = "BMI",
       y = "Frequency") +
  theme_minimal()
```
## BAR GRAPH TO VIEW THE SMOKING STATUS DISTRIBUTION

```{r}
# Assuming your column is named "smoking_status"
# Assuming your age column is named "age"

library(ggplot2)

# Filter data for age greater than or equal to 20
filtered_data <- stroke.data[stroke.data$age >= 20,]

ggplot(filtered_data, aes(x = factor(smoking_status))) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Smoking Status Distribution (Age >= 20)",
       x = "Smoking Status",
       y = "Frequency") +
  theme_minimal()

```
## BAR GRAPH TO VIEW THE SMOKING STATUS DISTRIBUTION IN STROKE  POSITIVE PATIENTS

```{r}
# Assuming your columns are named "smoking_status" and "stroke"
library(ggplot2)

ggplot(subset(stroke.data, stroke == 1), aes(x = factor(smoking_status))) +
  geom_bar(fill = "coral", color = "black") +
  labs(title = "Smoking Status for Stroke-Positive Patients",
       x = "Smoking Status",
       y = "Frequency") +
  theme_minimal()
```
## HISTOGRAM TO VIEW THE AGE DISTRIBUTION

```{r}
hist(stroke.data$age)
```
## BOXPLOT TO VIEW THE AGE DISTRIBUTION IN STROKE POSITIVE PATIENTS

```{r}
# Assuming your columns are named "age" and "stroke"
library(ggplot2)

ggplot(subset(stroke.data, stroke == 1), aes(x = factor(stroke), y = age)) +
  geom_boxplot(fill = "lightgreen", color = "black") +
  labs(title = "Distribution of Age for Stroke-Positive Patients",
       x = "Stroke",
       y = "Age") +
  theme_minimal()
```
## BAR GRAPH TO VIEW THE GENDER DISTRIBUTION
```{r}
library(ggplot2)
ggplot(stroke.data, aes(x = factor(gender))) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Distribution of Gender", x = "Gender", y = "Frequency") +
  theme_minimal()
```
## BAR GRAPH TO VIEW THE GENDER DISTRIBUTION IN STROKE POSITIVE PATIENTS
```{r}
library(ggplot2)
ggplot(subset(stroke.data, stroke == 1), aes(x = factor(stroke), fill = gender)) +
  geom_bar(position = "dodge", color = "black") +
  labs(title = "Distribution of Gender for Stroke-Positive Patients",
       x = "Stroke",
       y = "Frequency",
       fill = "Gender") +
  theme_minimal()
```
## statistical analysis for research questions

## Question1

```{r}
set.seed(123)  # Setting a seed for reproducibility
sample_size <- min(5000, nrow(stroke.data))  
sample_data <- stroke.data[sample(1:nrow(stroke.data), sample_size), ]
# Performing Shapiro-Wilk test
shapiro_test_result <- shapiro.test(sample_data$bmi)
# Printing the result
print(shapiro_test_result)
# Checking if the p-value is less than 0.05
if (shapiro_test_result$p.value < 0.05) {
  cat("The BMI data is not normally distributed based on the Shapiro-Wilk test.\n")
} else {
  cat("The BMI data appears to be normally distributed based on the Shapiro-Wilk test.\n")
}
```

```{r}
# Assuming your data frame is named stroke.data and "BMI" and "stroke" are the columns of interest
wilcoxon_test_result <- wilcox.test(bmi ~ stroke, data = stroke.data)

# Print the result
print(wilcoxon_test_result)

# Check if the p-value is less than 0.05
if (wilcoxon_test_result$p.value < 0.05) {
  cat("There is a significant difference in BMI between stroke-affected and non-affected individuals.There is an association between BMI and stroke occurrence\n")
} else {
  cat("There is no significant difference in BMI between stroke-affected and non-affected individuals.There is no association between BMI and stroke occurrence\n")
}

```
## Question2
```{r}
# Random Sampling
set.seed(123)  # Setting a seed for reproducibility
sample_size <- min(5000, nrow(stroke.data))  
sample_data <- stroke.data[sample(1:nrow(stroke.data), sample_size), ]
# Excluding samples with age less than 20
sample_data_filtered <- sample_data[sample_data$age >= 20, ]
# Creating a contingency table
cross_table <- table(sample_data_filtered$smoking_status, sample_data_filtered$stroke)
# Performing Chi-squared test
chi_square_test <- chisq.test(cross_table)
# Printing the result
#print(cross_table)
print(chi_square_test)
# Make an inference based on the p-value
if (chi_square_test$p.value < 0.05) {
  cat("There is a significant association between smoking status and stroke risk.\n")
} else {
  cat("There is no significant association between smoking status and stroke risk.\n")
}
```
## Question3

## age
```{r}
# Random Sampling
set.seed(123)  # Setting a seed for reproducibility
sample_size <- min(5000, nrow(stroke.data)) 
sample_data <- stroke.data[sample(1:nrow(stroke.data), sample_size), ]
# Performing Shapiro-Wilk test
shapiro_test_result <- shapiro.test(sample_data$age)
# Printing the result
print(shapiro_test_result)
# Checking if the p-value is less than 0.05
if (shapiro_test_result$p.value < 0.05) {
  cat("The Age data is not normally distributed based on the Shapiro-Wilk test.\n")
} else {
  cat("The Age data appears to be normally distributed based on the Shapiro-Wilk test.\n")
}
```
```{r}
# Random Sampling
set.seed(123)  # Setting a seed for reproducibility
sample_size <- min(5000, nrow(stroke.data))  
sample_data <- stroke.data[sample(1:nrow(stroke.data), sample_size), ]
# Performing Wilcoxon rank-sum test
wilcox_test_result <- wilcox.test(age ~ stroke, data = sample_data)
# Printing the result
print(wilcox_test_result)
# Printing association based on p-value
if (wilcox_test_result$p.value < 0.05) {
  cat("There is a significant association between age and stroke.\n")
} else {
  cat("There is no significant association between age and stroke.\n")
}
```
## gender

```{r}
set.seed(123)  # Setting a seed for reproducibility
sample_size <- min(5000, nrow(stroke.data))  
sample_data <- stroke.data[sample(1:nrow(stroke.data), sample_size), ]
contingency_table <- table(sample_data$gender, sample_data$stroke)
chi_square_test <- chisq.test(contingency_table)
print(chi_square_test)
if (chi_square_test$p.value < 0.05) {
  cat("There is a significant association between gender and stroke risk.\n")
} else {
  cat("There is no significant association between gender and stroke risk.\n")
}
```
## logistic regression

```{r}
# Random Sampling
set.seed(123)  # Setting a seed for reproducibility
sample_size <- min(5000, nrow(stroke.data))  # Ensure the sample size is within the valid range
sample_data <- stroke.data[sample(1:nrow(stroke.data), sample_size), ]

# Performing Logistic Regression
logistic_model <- glm(stroke ~ bmi, data = sample_data, family = binomial)

# Printing the model summary
summary(logistic_model)

```

```{r}
# Random Sampling
set.seed(123)  # Setting a seed for reproducibility
sample_size <- min(5000, nrow(stroke.data))  
sample_data <- stroke.data[sample(1:nrow(stroke.data), sample_size), ]
# Performing Logistic Regression
logistic_model <- glm(stroke ~ smoking_status, data = sample_data, family = binomial)
# Printing the model summary
summary(logistic_model)
```

```{r}
# Random Sampling
set.seed(123)  # Setting a seed for reproducibility
sample_size <- min(5000, nrow(stroke.data))  
sample_data <- stroke.data[sample(1:nrow(stroke.data), sample_size), ]
# Performing Logistic Regression
logistic_model_age <- glm(stroke ~ age, data = sample_data, family = binomial)
# Printing the model summary
summary(logistic_model_age)
```

```{r}
# Random Sampling
set.seed(123)  # Setting a seed for reproducibility
sample_size <- min(5000, nrow(stroke.data)) 
sample_data <- stroke.data[sample(1:nrow(stroke.data), sample_size), ]
# Performing Logistic Regression
logistic_model_gender <- glm(stroke ~ gender, data = sample_data, family = binomial)
# Printing the model summary
summary(logistic_model_gender)
```
```{r}
# Random Sampling
set.seed(123)  # Setting a seed for reproducibility
sample_size <- min(5000, nrow(stroke.data))
sample_data <- stroke.data[sample(1:nrow(stroke.data), sample_size), ]

# Exclude samples with age less than 20 for smoking_status
filtered_smoking_status <- subset(sample_data, !(age < 20 & !is.na(smoking_status)))

# Perform Logistic Regression with filtered smoking_status
logistic_model_combined <- glm(stroke ~ bmi + smoking_status + age + gender, data = filtered_smoking_status, family = binomial)

# Printing the model summary
summary(logistic_model_combined)
```


